Measurement and Probability in Fuzzy Quantum Theories

### 2016-11-28 proposal/prop.tex
hanson

\begin{figure}
\begin{center}
\includegraphics[scale=0.48]{measure4.pdf}
\end{center}
\caption{\label{fig:three-dimensional-4-value}This figure illustrates
  cases 1 and 2 of example~\ref{ex:three-dimensional-4-value} plotted
  in $\mathbb{R}^{3}$. The red and green dotted vectors are $\ket{0}$
  and $\ket{\ps}$ respectively.  All possible vectors of the subspaces
  $\ket{0_{\theta,\gamma}^{\perp}}$ and
  $\ket{\ps_{\theta,\gamma}^{\perp}}$ are drawn in the red and green
  circles, respectively. Within the circles, a dotted
  vector~$\ket{\psi}$ means $\bar{\mu}(\proj{\psi})=\likely$;
  otherwise, $\bar{\mu}(\proj{\psi})=\unlikely$. The gray vector is a
  generic vector~$\ket{0_{\theta,\gamma}^{\perp}}$, and the red and
  green solid vectors are normalized
  $\ket{0}\times\ket{0_{\theta,\gamma}^{\perp}}$ and
  $\ket{\ps}\times\ket{0_{\theta,\gamma}^{\perp}}$, respectively.}
\end{figure}

@@

\begin{figure}
\begin{center}
  \includegraphics[scale=0.48]{HoughLineData.pdf}
\end{center}
  \caption{Curves of possible line parameters determined by noisy
    data samples on a straight line in 2D space.}
\label{houghline.fig}
\end{figure}

@@

\noindent (A) {\em Integration of research and education:} Courses on
advanced computational models and quantum information will be further
developed at Indiana University to communicate the basic concepts and
tools discussed in this proposal. Dr. Ortiz taught courses on
``Quantum information and Computation'' and Dr. Sabry has taught
courses on ``Quantum Computing'' for graduate students.  These courses
were highly appreciated by the students as noted by their evaluations
in making the material extremely accessible and interesting.  The PIs
plan to engage high school students or undergraduates in the
implementation of some of the ideas.  The PIs have a long standing
commitment to K-12 science and math education (see also (C)
below). The collaborative group efforts will provide students, as well
as the PIs with a unique learning experience. The broadened
perspectives gained through those efforts will contribute to the PIs
effectiveness as teachers and mentors. They also will enable to better
prepare students and postdocs for productive careers in an
increasingly competitive global economy. Students and postdocs are
encouraged to pursue collaborative research leading to a broader
scientific program. This is highlighted by the PIs students' close
contact with LANL, the Perimeter Institute, and the Institute for
Quantum Computing at Waterloo.

\medskip
\noindent (B) {\em Broadening the participation of underrepresented groups:}
Indiana University is engaged in attracting women and minorities in all
ranks. Several programs are aimed at specifically attracting outstanding
students into physics and computer science, and to expanding the pool of
candidates for fellowships. Targeted recruitment of minority graduate
students is facilitated by the Falicov Fellowship established in 2006. In
recent years, the School of Informatics and Computing has been fortunate to
hire several women and minorities for faculty positions, and we have been
working toward increasing the number of women and minorities among our
student population. To assist with this process, the PIs will encourage women
and minorities to apply for graduate positions as well as undergraduate
positions that may be made available by the grant and REU
supplements. Dr. Ortiz has minority students and Dr. Sabry has graduated a
female Ph.D. in 2005 and is currently assisting on the Ph.D. committees of
several others. Efforts will be made to broaden diversity.

@@

\paragraph*{Prior Support.} There was no recent NSF support for
Dr. Ortiz. In the past five years, Prof. Sabry received one NSF grant
(award number 1217454) titled \emph{Information Effects} for the
amount of \$275,319.00 with start and end dates of 1 September 2012 --
31 August 2016.

@@

\paragraph*{Publications.} Several papers were published to report on
the results. On the foundational side, we have the following papers:
(1) Computing with Semirings and Weak Rig Groupoids, European
Symposium on Programming, 2016, (2) Superstructural Reversible Logic,
International Workshop on Linearity, 2014, (3) Theseus: A High Level
Language for Reversible Computing. Work-in-Progress report in the
Conference on Reversible Computation, 2014, (4) Discrete Quantum
Theories, J. Phys. A: Math. Theor. 2014., (5) Geometry of Discrete
Quantum Computing, J. Phys. A: Math. Theor. 2013., (6) Isomorphic
Interpreters from Logically Reversible Abstract Machines. Workshop on
Reversible Computation 2012. On the applications side, we have the
following papers: (1) Copredication in Homotopy Type Theory: A
homotopical approach for formal semantics of natural
languages. Workshop on Natural Language and Computer Science, 2016.,
(2) Contract Monitoring Semantics as Patterns of
Communication. International Conference on Functional Programming,
2015., (3) Extensible Effects: An Alternative to Monad
Transformers. Haskell Symposium, 2013., (4) Encoding Secure
Information Flow with Restricted Delegation and Revocation in
Haskell. Workshop on Functional Programming Concepts in
Domain-Specific Languages, 2013. In addition to the publications,
Dr. Sabry and collaborators maintain a public repository with software
related to the project. Dr. Sabry was also invited to give talks at
several venues, most notably the Colloquiumfest, University of
Saskatchewan, Canada, 2014, and the Workshop on Quantum Information
and Foundations of Quantum Mechanics, University of British Columbia,
Vancouver, Canada, 2013

>>>>

\begin{figure}
\begin{center}
\includegraphics[scale=0.38]{measure4.pdf}
\end{center}
\caption{\label{fig:three-dimensional-4-value}This figure illustrates
  cases 1 and 2 of example~\ref{ex:three-dimensional-4-value} plotted
  in $\mathbb{R}^{3}$. The red and green dotted vectors are $\ket{0}$
  and $\ket{\ps}$ respectively.  All possible vectors of the subspaces
  $\ket{0_{\theta,\gamma}^{\perp}}$ and
  $\ket{\ps_{\theta,\gamma}^{\perp}}$ are drawn in the red and green
  circles, respectively. Within the circles, a dotted
  vector~$\ket{\psi}$ means $\bar{\mu}(\proj{\psi})=\likely$;
  otherwise, $\bar{\mu}(\proj{\psi})=\unlikely$. The gray vector is a
  generic vector~$\ket{0_{\theta,\gamma}^{\perp}}$, and the red and
  green solid vectors are normalized
  $\ket{0}\times\ket{0_{\theta,\gamma}^{\perp}}$ and
  $\ket{\ps}\times\ket{0_{\theta,\gamma}^{\perp}}$, respectively.}
\end{figure}

@@

\begin{figure}
\begin{center}
  \includegraphics[scale=0.38]{HoughLineData.pdf}
\end{center}
  \caption{Curves of possible line parameters determined by noisy
    data samples on a straight line in 2D space.}
\label{houghline.fig}
\end{figure}

@@

\noindent (A) {\em Integration of research and education:} Courses on
advanced computational models and quantum information will be further
developed at Indiana University to communicate the basic concepts and
tools discussed in this proposal. Dr. Ortiz taught courses on
``Quantum information and Computation'' and Dr. Sabry has taught
courses on ``Quantum Computing'' for graduate students.  Dr.\ Hanson
has taught courses on scientific visualization and mathematical
modeling with emphasis on high-dimensional mathematics.  His numerous
Siggraph course presentations are embodied in his 2006 book {\it
  Visualizing Quaternions\/}, which remains the definitive treatment
of the subject for computer science and data analysis applications and
has sold thousands of copies. As a part of Dr.\ Hanson's commitment to
exposing and explaining the elements of higher mathematics to the
public, he has recently led the design of a free iPhone App, {\it
  4Dice\/}, that for the first time makes available an interactive
treatment of a correctly imaged simulation of the appearance of a
hypercube, realizing interactively his YouTube animation of the same
name.  This framework is now being generalized for a broad family of
iPhone Apps to permit the hands-on study of high-dimensional
mathematics.  These courses were highly appreciated by the students as
noted by their evaluations in making the material extremely accessible
and interesting.  The PIs plan to engage high school students or
undergraduates in the implementation of some of the ideas.  The PIs
have a long standing commitment to K-12 science and math education
(see also (C) below). The collaborative group efforts will provide
students, as well as the PIs with a unique learning experience. The
broadened perspectives gained through those efforts will contribute to
the PIs effectiveness as teachers and mentors. They also will enable
to better prepare students and postdocs for productive careers in an
increasingly competitive global economy. Students and postdocs are
encouraged to pursue collaborative research leading to a broader
scientific program. This is highlighted by the PIs students' close
contact with LANL, the Perimeter Institute, and the Institute for
Quantum Computing at Waterloo.

\medskip
\noindent (B) {\em Broadening the participation of underrepresented groups:}
Indiana University is engaged in attracting women and minorities in all
ranks. Several programs are aimed at specifically attracting outstanding
students into physics and computer science, and to expanding the pool of
candidates for fellowships. Targeted recruitment of minority graduate
students is facilitated by the Falicov Fellowship established in 2006. In
recent years, the School of Informatics and Computing has been fortunate to
hire several women and minorities for faculty positions, and we have been
working toward increasing the number of women and minorities among our
student population. To assist with this process, the PIs will encourage women
and minorities to apply for graduate positions as well as undergraduate
positions that may be made available by the grant and REU
supplements. Dr. Ortiz has minority students and Dr. Sabry has graduated a
female Ph.D. in 2005 and is currently assisting on the Ph.D. committees of
several others.

@@

\paragraph*{Prior Support.} There was no recent NSF support for
Dr. Ortiz or Dr. Hanson. In the past five years, Prof. Sabry received
one NSF grant (award number 1217454) titled \emph{Information Effects}
for the amount of \$275,319.00 with start and end dates of 1 September
2012 -- 31 August 2016.

@@

\paragraph*{Publications.} Several papers were published to report on
the results. On the foundational side, we have the following papers:
(1) Computing with Semirings and Weak Rig Groupoids, ESOP, 2016, (2)
Superstructural Reversible Logic, Workshop on Linearity, 2014, (3)
Theseus: A High Level Language for Reversible
Computing. Work-in-Progress report, RC 2014, (4) Discrete Quantum
Theories, J. Phys. A: Math. Theor. 2014., (5) Geometry of Discrete
Quantum Computing, J. Phys. A: Math. Theor. 2013., (6) Isomorphic
Interpreters from Logically Reversible Abstract Machines. RC 2012. On
the applications side, we have the following papers: (1) Copredication
in Homotopy Type Theory: A homotopical approach for formal semantics
of natural languages. Workshop on Natural Language and Computer
Science, 2016., (2) Contract Monitoring Semantics as Patterns of
Communication. ICFP, 2015., (3) Extensible Effects: An Alternative to
Monad Transformers. Haskell Symposium, 2013., (4) Encoding Secure
Information Flow with Restricted Delegation and Revocation in
Haskell. Workshop on Functional Programming Concepts in
Domain-Specific Languages, 2013. In addition to the publications,
Dr. Sabry and collaborators maintain a public repository with software
related to the project. Dr. Sabry was also invited to give talks at
several venues, most notably the Colloquiumfest, University of
Saskatchewan, Canada, 2014, and the Workshop on Quantum Information
and Foundations of Quantum Mechanics, University of British Columbia,
Vancouver, Canada, 2013
